---
title: "QA_Module11"
author: "Nidhi Lovanshi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***Load Package***
```{r}
library(lpSolveAPI)
```

***Create a model***
```{r}

# Creating a model with 7 variables
model <- make.lp(0, 7)  

```

***Objective function coefficients***
```{r}
# Set the objective function coefficients
# Objective: minimize total wage
set.objfn(model, c(775, 800, 800, 800, 800, 775, 750))

# Required number of workers per day
workers_required <- c(18, 27, 22, 26, 25, 21, 19)


```

***Constraints***

```{r message=FALSE, warning=FALSE}
# Add constraints for each day of the week (corrected coefficients)

# Sunday: Shifts 2, 3, 4, 5, 6 are available
add.constraint(model, c(0, 1, 1, 1, 1, 1, 0), ">=", workers_required[1])

# Monday: Shifts 3, 4, 5, 6, 7 are available
add.constraint(model, c(0, 0, 1, 1, 1, 1, 1), ">=", workers_required[2])  

# Tuesday: Shifts 1, 4, 5, 6, 7 are available
add.constraint(model, c(1, 0, 0, 1, 1, 1, 1), ">=", workers_required[3])  

# Wednesday: Shifts 1, 2, 5, 6, 7 are available
add.constraint(model, c(1, 1, 0, 0, 1, 1, 1), ">=", workers_required[4])  

# Thursday: Shifts 1, 2, 3, 6, 7 are available
add.constraint(model, c(1, 1, 0, 0, 1, 1, 1), ">=", workers_required[5])  

# Friday: Shifts 1, 2, 3, 4, 7 are available
add.constraint(model, c(1, 1, 1, 1, 0, 0, 1), ">=", workers_required[6])  

# Saturday: Shifts 1, 2, 3, 5 are available
add.constraint(model, c(1, 1, 1, 0, 1, 0, 0), ">=", workers_required[7])  

# Set all variables to be integer
set.type(model, 1:7, "integer")

# # Set the problem to maximize minimize
lp.control(model, sense = "min")

# Solve the model
solve(model)
```

***Objective and Decision variables value***
```{r}
# Get the results
solution <- get.variables(model)
obj_value <- get.objective(model)
```

***Print the results***
```{r}
# Print the results
cat("Optimal solution found\n")
cat("Total cost: $", obj_value, "\n")
```
The optimal solution minimizes the total wage cost to $24,175
```{r}
cat("\nNumber of workers for each shift:\n")
for (i in 1:7) {
  cat("Shift", i, ":", solution[i],"workers", "\n")
}

```
Given the shifts and their days off, the distribution of workers among the seven shifts is as follows:

Shift 1 (Sunday and Monday): 1 workers  
Shift 2 (Monday and Tuesday): 3 workers  
Shift 3 (Tuesday and Wednesday): 5 workers   
Shift 4 (Wednesday and Thursday): 0 workers   
Shift 5 (Thursday and Friday): 10 workers  
Shift 6 (Friday and Saturday): 0 workers   
Shift 7 (Saturday and Sunday): 12 workers  

```{r}
# Calculating and printing workers available each day
days <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")

workers_available <- c(
  sum(solution[c(2, 3, 4, 5, 6)]),  # Sunday: Shifts 2, 3, 4, 5, 6 are available
  sum(solution[c(3, 4, 5, 6, 7)]),  # Monday: Shifts 3, 4, 5, 6, 7 are available
  sum(solution[c(1, 4, 5, 6, 7)]),  # Tuesday: Shifts 1, 4, 5, 6, 7 are available
  sum(solution[c(1, 2, 5, 6, 7)]),  # Wednesday: Shifts 1, 2, 5, 6, 7 are available
  sum(solution[c(1, 2, 3, 6, 7)]),  # Thursday: Shifts 1, 2, 3, 6, 7 are available
  sum(solution[c(1, 2, 3, 4, 7)]),  # Friday: Shifts 1, 2, 3, 4, 7 are available
  sum(solution[c(1, 2, 3, 5)])      # Saturday: Shifts 1, 2, 3, 5 are available
)

# Print the available workers for each day
cat("\nWorkers available each day:\n")
for (i in 1:7) {
  cat(days[i], ": ", workers_available[i],  "\n", sep="")
}

```

