---
title: "Assignment4 - Module 4 - Solve LP Model"
author: "Nidhi Lovanshi"
date: "`r Sys.Date()`"
output: html_document
---
**Load necessary Packages**

```{r message=FALSE, warning=FALSE}
library(lpSolveAPI)

```

**Reading lp file**

```{r message=FALSE, warning=FALSE}
# We have created a lp file which contains
lp_model <-read.lp("qmm.lp")
```

**Displaying the lp object "lp_model"**
```{r message=FALSE, warning=FALSE}
print(lp_model)
summary(lp_model)
```

**Solving the Linear programming model**
```{r message=FALSE, warning=FALSE}
solve(lp_model)
```


0 indicates this has ran successfully and there is an optimal solution 

```{r}
#Objective Function:
get.objective(lp_model)
```
The objective function max Z is 420 x11 + 360 x12 + 300 x13 + 420 y11 + 360 y12 + 300 y13 + 420 z11 + 360 z12 + 300 z13

Here, the optimal value for Z is $696000.


```{r}
#Variables:
get.variables(lp_model)
```

The optimal value for each variables to obtain above optimal solution are:  
x11 = 516.6667  
x12 = 177.7778  
x13 = 0.0000  
y11 = 0.0000  
y12 = 666.6667  
y13 = 166.6667  
z11 = 0.0000  
z12 = 0.0000  
z13 = 416.6667  

1) Plant 1 should produce 516.67 large products and 177.78 medium products but no small products.
2) Plant 2 should produce 0 large products,666.67 medium and 166.67 small products
3) Plant 3 should produce 0 large products, 0 medium and 416.67 small products

```{r}
#Constraints:
get.constraints(lp_model)
```


All the constraints value are satisfied 



Conclusion, the optimal maximum profit Z is $696000 with all the above optimal variables values with all satisfied constraints. 