---
title: "Module 6 - The Transportation Model"
author: "Nidhi Lovanshi"
date: "`r Sys.Date()`"
output: html_document
---
  **Load necessary Packages**
  
```{r message=FALSE, warning=FALSE}
library(lpSolveAPI)

```

**Reading lp file**
  
```{r message=FALSE, warning=FALSE}
# We have created a lp file which contains
lp_model <-read.lp("qmm2.lp")
```

**Displaying the lp object "lp_model"**
```{r message=FALSE, warning=FALSE}
print(lp_model)
summary(lp_model)
```

**Solving the Linear programming model**
```{r message=FALSE, warning=FALSE}
solve(lp_model)
```


0 indicates this has ran successfully and there is an optimal solution 


```{r}
#Objective Function:
get.objective(lp_model)
```

The objective function min Z is 600 x11 + 600 x12 + 600 x13 + 625 y11 + 625 y12 + 625 y13 + 22 x11 + 14 x12 + 30 x13 + 16 y11 + 20 y12 + 24 y13

Here, the optimal value for Z is $ 132790.


```{r}
#Variables:
get.variables(lp_model)
```


The optimal value for each variables to obtain above optimal solution are:  
x11 = 0   
x12 = 60    
x13 = 40   
y11 = 80    
y12 = 0    
y13 = 30   

1) No units are shipped from Plant A to Warehouse 1.
2) 60 units are shipped from Plant A to Warehouse 2.
3) 40 units are shipped from Plant A to Warehouse 3.
4) 80 units are shipped from Plant B to Warehouse 1.
5) No units are shipped from Plant B to Warehouse 2.
6) 30 units are shipped from Plant B to Warehouse 3. 

```{r}
#Constraints:
get.constraints(lp_model)
```

All the constraints value are satisfied 



Conclusion, the optimal minimum cost Z is $ 132790 with all the above optimal variables values with all satisfied constraints. 
