---
title: "QMM_FINAL"
author: "Nidhi Lovanshi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***Load necessary Packages***
```{r}
# Load library
library(lpSolveAPI)

```

***Generating Data***
```{r}
# Setting random seed for reproducibility
set.seed(123)

# Generating Dataset containing Student IDs, GPA, Skill Relevance and Leadership scores
data <- data.frame(
  ID = seq(1, 12), 
  GPA = runif(12, min = 2.0, max = 4.0),  
  SkillRelevance = runif(12, min = 0, max = 10), 
  Leadership = runif(12, min = 0, max = 10)  
)

```
A data frame is generated containing random information for 12 students.  
This includes their:  
ID: A unique number assigned to each student.  
GPA: Randomly assigned GPA scores ranging from 2.0 to 4.0.  
SkillRelevance: Random scores that indicate how relevant their skills are, on a scale from 0 to 10.  
Leadership: Random scores that reflect their leadership qualities, also on a scale from 0 to 10.  

***Assign weights***
```{r}
# Assigning weights to the factors contributing to success
weight_gpa <- 0.4
weight_skill <- 0.3
weight_leadership <- 0.3
```
These weights represent a careful strategy to maximize the group's performance, making sure that each member plays a unique role in the overall success.  

The reason for these weights is that GPA is considered more important, so it gets a higher weight of 0.4. On the other hand, Skill Relevance and Leadership are seen as equally important, so they both have a weight of 0.3.



***Student Success Score***
```{r}
#Computing each student's success score
data$SuccessScore <- with(data, weight_gpa * GPA + weight_skill * SkillRelevance + weight_leadership * Leadership)

```

Calculating a combined score (SuccessScore) for every student by merging their GPA, SkillRelevance, and Leadership scores, with each factor given a specific weight based on its significance.


***Configure the Linear Programming Model***
```{r}
#Parameters
total_students <- nrow(data)  
total_groups <- 4  
group_size <- 3  
```
Parameters: 

Total number of students (total_students) is 12.  
Total number of groups(total_groups) is 4.   
Number of students in each group(group_size) is 3.
```{r}
# Initializing the linear programming model
lp_model <- make.lp(0, total_students * total_groups)

# Defining the objective function
#The goal is to maximize the total SuccessScore across all groups.
objective <- rep(data$SuccessScore, times = total_groups)
set.objfn(lp_model, objective)

# Constraint : Each student should be assigned to one and only one group
# A loop goes through every student and sets a rule for their assignment.

for (student in 1:total_students) {
  constraint <- numeric(total_students * total_groups)
  for (group in 1:total_groups) {
    constraint[(group - 1) * total_students + student] <- 1
  }
  add.constraint(lp_model, constraint, "=", 1)
}

# Constraint : Each group should have exactly 3 members
# A loop goes through each group and sets a rule that the total number of students assigned to that group must be 3.

for (group in 1:total_groups) {
  constraint <- numeric(total_students * total_groups)
  group_indices <- ((group - 1) * total_students + 1):(group * total_students)
  constraint[group_indices] <- 1
  add.constraint(lp_model, constraint, "=", group_size)
}

# Specifying that all decision variables are binary
set.type(lp_model, columns = 1:(total_students * total_groups), type = "binary")
```
Each variable represents whether a student is assigned to a group (1) or not (0).  



***Solving the Linear programming model***
```{r}
solve(lp_model)

```
0 indicates this has ran successfully and there is an optimal solution


***Results and validation***
```{r}
#Variables:
variables <- get.variables(lp_model)
student_matrix <- matrix(variables, nrow = total_students, ncol = total_groups)

# Output results
cat("Assignment Matrix:\n")
print(student_matrix)
```
Each row represent a student (1 to 12). 
Each column represent a group (1 to 4).  


Column 1: Students 4, 8, 9 is assigned to Group 1.   
Column 2: Students 6, 10, 12 is assigned to Group 2.   
Column 3: Students 1, 2, 7 is assigned to Group 3.    
Column 4: Students 3, 5, 11 is assigned to Group 4.  

```{r}
# Verifying group sizes
group_sizes <- colSums(student_matrix)
cat("Group Sizes: ", group_sizes, "\n")

```

This confirms that each group contains exactly 3 students, satisfying the constraint.

```{r}
# Verifying each student is assigned to exactly one group
student_assignments <- rowSums(student_matrix)
cat("Student Assignments: ", student_assignments, "\n")
```
This confirms that every student is assigned to exactly one group, satisfying the constraint.


```{r}
# Determining which students are in each group
group_assignments <- lapply(1:total_groups, function(group) {
  which(student_matrix[, group] == 1)
})

# Displaying the group assignments and optimal success score
cat("Optimal Group Assignments:\n")
for (group in 1:total_groups) {
  cat(sprintf("Group %d: %s\n", group, paste(group_assignments[[group]], collapse = ", ")))
}
cat("\nOptimal Success Score: ", get.objective(lp_model), "\n")
```

The maximum total success score for the group assignments Z is 56.85392.  

The solution satisfies all constraints:
Each group contains exactly 3 students.
Each student is assigned to only one group.